The Poisson approximate likelihood compared to the particle filter

### 2024-07-24 ms.Rnw
compilation, dofuture

graphics.off()      # close graphics windows
library(pomp)
library(magrittr)
library(ggplot2)
library(scales)
library(foreach)
library(doParallel)
library(dplyr)
source("choose_n_fits.R")

RUN_LEVEL = 1
NP_MIF       = switch(RUN_LEVEL, 5000, 50000)
NMIF         = switch(RUN_LEVEL, 4,    100)
ncores       = switch(RUN_LEVEL, 4,    36)
top_n_fits   = switch(RUN_LEVEL, 1,    12)
@

@@

<<OvOv-model, include=False, massage=False, warning=False>>=
# measurement model 
dmeas <- Csnippet("
                  if (ISNA(cases1)) {
                  lik = (give_log) ? 0 : 1;
                  } else {
                        lik =  dbinom(cases1, H1, q1, 1) +
                        dbinom(cases2, H2, q2, 1) +
                        dbinom(cases3, H3, q3, 1);

@@

# define parameters (without betas)
params_fixed <- c(gamma=1, delta1=1/(5*52),delta2=1/(55*52), alpha=1/(78.86912*52), 
                  mu=0, N=82372825, omega=1/(1*52))
# WWR's rinit
rinit <- Csnippet("
    double m = N/(S10+I10+R10+S20+I20+R20+S30+I30+R30);
    I1=nearbyint(m*I10);
    I2=nearbyint(m*I20);
    I3=nearbyint(m*I30);
    S1=nearbyint(m*S10);
    S2=nearbyint(m*S20);
    S3=nearbyint(m*S30);
    R1=nearbyint(m*R10);
    R2=nearbyint(m*R20);
    R3=nearbyint(m*R30);
    H1 = 0;
    H2 = 0;
    H3 = 0;
")

@@

sir_panel <- panelPomp::panelPomp(list(unit1=sir),
                                  shared=NULL,
                                  specific=wwr_mle |> 
                                    as.matrix() |>
                                    `colnames<-`("unit1")
)
@

@@

<<maximization1, include=False, massage=False, warning=False>>=
set.seed(123)
sir_box <- rbind(
  beta1=c(10,15),
  beta2=c(0.2,0.4),
  beta3=c(0.3,0.5),
  phi=c(0.01,0.3),
  beta11=c(0.1,0.2),
  sigma_q=c(0.001,0.1),
  sigma_xi=c(65,70),
  S10 = c(0.01,0.05),
  I10 = c(0.0001, 0.0005),
  R10 = c(0.01, 0.05),
  S20 = c(0.01,0.7),
  I20 = c(0.000001,0.00002),
  R20 = c(0.001,0.005),
  S30 = c(0.01,0.5),
  I30 = c(0.000001, 0.0001),
  R30 = c(0.0001,0.005)
)

## Make it panelPomp
c(apply(sir_box,1,function(x)runif(1,min=x[1],max=x[2])),
  params_fixed) |> 
  as.matrix() |>
  `colnames<-`("unit1") -> starting

require(doParallel)
registerDoParallel(ncores)
mcopts <- list(set.seed=TRUE)

bake(file = "output/round_01/ovov_mif_01.rds",{ 
  mifs_global_01 <- foreach(i=1:ncores,.packages='pomp', .options.multicore=mcopts) %dopar% {
    set.seed(112*i)
    panelPomp::mif2(
      sir_panel,
      Np = NP_MIF,
      cooling.fraction.50 = 0.5,
      rw.sd = rw_sd(beta1=0.01,beta2=0.01,beta3=0.01,
                    beta11=0.01,phi=0.01,sigma_q=0.01,sigma_xi=0.01,
                    S10=ivp(0.24),I10=ivp(0.24),R10=ivp(0.24),
                    S20=ivp(0.24),I20=ivp(0.24),R20=ivp(0.24),
                    S30=ivp(0.24),I30=ivp(0.24),R30=ivp(0.24)),
      cooling.type = "geometric",
      Nmif = NMIF,
      shared.start = numeric(0),
      specific.start = starting,
      block = F
    ) 
  }
  mifs_global_01
})

bake(file = "output/round_01/ovov_01_el.rds",{
  el_01 <- measlespkg::eval_logLik(mifs_global_01, ncores=ncores, np_pf = NP_MIF, nreps=ncores)
  el_01
})
@

@@

<<maximization2, include=False, massage=False, warning=False>>=
set.seed(1234)
sd_02           = 2/3

el_01 <- readRDS("output/round_01/ovov_01_el.rds")
starting_values_02 <- choose_n_fits(el_01, top_n_fits, ncores)

require(doParallel)
registerDoParallel(ncores)
mcopts <- list(set.seed=TRUE)

bake(file = "output/round_02/ovov_mif_02.rds",{ 
  mifs_global_02 <- foreach(i=1:ncores,.packages='pomp', .options.multicore=mcopts) %dopar% {
  panelPomp::mif2(
    sir_panel,
    Np = NP_MIF,
    cooling.fraction.50 = 0.5,
    rw.sd = rw_sd(beta1=0.01*sd_02,beta2=0.01*sd_02,beta3=0.01*sd_02,
                  beta11=0.01*sd_02,phi=0.01*sd_02,sigma_q=0.01*sd_02,sigma_xi=0.01*sd_02,
                  S10=ivp(0.24)*sd_02,I10=ivp(0.24)*sd_02,R10=ivp(0.24)*sd_02,
                  S20=ivp(0.24)*sd_02,I20=ivp(0.24)*sd_02,R20=ivp(0.24)*sd_02,
                  S30=ivp(0.24)*sd_02,I30=ivp(0.24)*sd_02,R30=ivp(0.24)*sd_02),
    cooling.type = "geometric",
    Nmif = NMIF,
    shared.start = numeric(0),
    specific.start = starting_values_02[[i]],
    block = F
  ) 
}
  mifs_global_02
})

bake(file = "output/round_02/ovov_02_el.rds",{
  el_02 <- measlespkg::eval_logLik(mifs_global_02, ncores = ncores, np_pf = NP_MIF, nreps=ncores)
  el_02
})
@

@@

<<maximization3, include=False, massage=False, warning=False>>=
set.seed(12345)
sd_03           = (2/3)^2

el_02 <- readRDS("output/round_02/ovov_02_el.rds")
starting_values_03 <- choose_n_fits(el_02, top_n_fits, ncores)

require(doParallel)
registerDoParallel(ncores)
mcopts <- list(set.seed=TRUE)

bake(file = "output/round_03/ovov_mif_03.rds",{ 
  mifs_global_03 <- foreach(i=1:ncores,.packages='pomp', .options.multicore=mcopts) %dopar% {
  panelPomp::mif2(
    sir_panel,
    Np = NP_MIF,
    cooling.fraction.50 = 0.5,
    rw.sd = rw_sd(beta1=0.01*sd_03,beta2=0.01*sd_03,beta3=0.01*sd_03,
                  beta11=0.01*sd_03,phi=0.01*sd_03,sigma_q=0.01*sd_03,sigma_xi=0.01*sd_03,
                  S10=ivp(0.24)*sd_03,I10=ivp(0.24)*sd_03,R10=ivp(0.24)*sd_03,
                  S20=ivp(0.24)*sd_03,I20=ivp(0.24)*sd_03,R20=ivp(0.24)*sd_03,
                  S30=ivp(0.24)*sd_03,I30=ivp(0.24)*sd_03,R30=ivp(0.24)*sd_03),
    cooling.type = "geometric",
    Nmif = NMIF,
    shared.start = numeric(0),
    specific.start = starting_values_03[[i]],
    block = F
  ) 
}
  mifs_global_03
})

bake(file = "output/round_03/ovov_03_el.rds",{
  el_03 <- measlespkg::eval_logLik(mifs_global_03, ncores = ncores, np_pf = NP_MIF, nreps=ncores)
  el_03
})
@

<<estimated-mle, include=False, massage=False, warning=False>>=
mif_el <- readRDS("output/round_03/ovov_03_el.rds") # Can be changed if more rounds needed
est_mle_vec <- mif_el$fits[which.max(mif_el$fits$logLik),]
@

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.55\textwidth]{thesis-figs/PAL_vs_SMC_ovov_100.png}
\caption{\label{fig:PAL_vs_SMC_ovov_100}Log-likelihoods computed using two filtering methods for 100 randomly simulated datasets at the MLE of the OvOv model.
The particle filter and PAL used 50,000 particles.
The red line is the line $Lik(pfilter)=Lik(palfilter)$.
On average, the particle filter likelihood is 7.7 log units higher than the PAL.
%%  -6224.245 - (-6231.97) = 7.725
}
\end{figure}

@@

\begin{table}[htbp]
  \centering
  \begin{threeparttable}
    \caption{Maximum likelihood estimation of parameters by PAL for three models \citep{wwr}, and by mif2 algorithm \citep{pomppackagepaper} for the OvOv model.
    \ed{LET'S DISCUSS WHETHER WE WANT TO SHOW JUST OV-OV HERE. IT WOULD BE SIMPLER.}
    }
    \label{tab:mlebywwr}
    \begin{tabular}{>{\raggedright\arraybackslash}p{2cm} *{3}{>{\centering\arraybackslash}p{1.5cm}} >{\centering\arraybackslash}p{2.5cm}}
      \toprule
      Parameter      & EqEq  & EqOv  & OvOv  & OvOv (\textbf{pomp})$^*$ \\
      \midrule
      \(\beta_1\)    & 12.15 & 12.74 & 11.48 & $\Sexpr{est_mle_vec["beta1[unit1]"]}$\\
      \(\beta_2\)    & 0.22  & 0.21  & 0.25  & $\Sexpr{est_mle_vec["beta2[unit1]"]}$\\
      \(\beta_3\)    & 0.34  & 0.31  & 0.35  & $\Sexpr{est_mle_vec["beta3[unit1]"]}$\\
      \(\phi\)       & 0.017 & 0.14  & 0.14  & $\Sexpr{est_mle_vec["phi[unit1]"]}$\\
      \(\rho\)       & 0.022 & 0.19  & 0.16  & $\Sexpr{est_mle_vec["rho[unit1]"]}$\\
      \(\sigma^2_q\) & n/a   & 0.042 & 0.021 & $\Sexpr{est_mle_vec["sigma_q[unit1]"]}$\\
      \(\sigma_\xi\) & n/a   & n/a   & 66.89 & $\Sexpr{est_mle_vec["sigma_xi[unit1]"]}$\\
      \addlinespace
      \multicolumn{1}{l}{\(S_{10}\)} & \multicolumn{3}{c}{$3876549^*$} & $$\Sexpr{est_mle_vec["S10[unit1]"]}$$\\
      \multicolumn{1}{l}{\(I_{10}\)} & \multicolumn{3}{c}{$30351$}     & $$\Sexpr{est_mle_vec["I10[unit1]"]}$$\\
      \multicolumn{1}{l}{\(R_{10}\)} & \multicolumn{3}{c}{$1315221$}   & $$\Sexpr{est_mle_vec["R10[unit1]"]}$$\\
      \multicolumn{1}{l}{\(S_{20}\)} & \multicolumn{3}{c}{$57139612$}  & $$\Sexpr{est_mle_vec["S20[unit1]"]}$$\\
      \multicolumn{1}{l}{\(I_{20}\)} & \multicolumn{3}{c}{$871$}       & $$\Sexpr{est_mle_vec["I20[unit1]"]}$$\\
      \multicolumn{1}{l}{\(R_{20}\)} & \multicolumn{3}{c}{$302852$}    & $$\Sexpr{est_mle_vec["R20[unit1]"]}$$\\
      \multicolumn{1}{l}{\(S_{30}\)} & \multicolumn{3}{c}{$19573727$}  & $$\Sexpr{est_mle_vec["S30[unit1]"]}$$\\
      \multicolumn{1}{l}{\(I_{30}\)} & \multicolumn{3}{c}{$2550$}      & $$\Sexpr{est_mle_vec["I30[unit1]"]}$$\\
      \multicolumn{1}{l}{\(R_{30}\)} & \multicolumn{3}{c}{$131092$}    & $$\Sexpr{est_mle_vec["R30[unit1]"]}$$\\
      \midrule
      \addlinespace
      \textbf{AIC}  & $98866.65$ & $15154.75$ & $13778.08$ & $\Sexpr{(-2)*((length(wwr_mle) - length(params_fixed)) - est_mle_vec["logLik"])}$\\
      \addlinespace
      \bottomrule
    \end{tabular}
    \begin{tablenotes}
      \small
      \item[*] The initial distribution parameter \(\lambda_0 = (S_{10}, I_{10}, R_{10}, S_{20}, I_{20}, R_{20}, S_{30}, I_{30}, R_{30})\) are assumed to be \textbf{fixed} in \cite{wwr}. The results of the estimates of the OvOv (\textbf{pomp}) model are obtained by using the Iterated Filtering algorithm with 3 rounds and 100 iterations, 50,000 particles, and 36 replicates in each round with the top 12 best fits in terms of likelihood chosen to be the starting value for the next round. 
    \end{tablenotes}
  \end{threeparttable}
\end{table}

>>>>

graphics.off()      # close graphics windows
library(pomp)
library(magrittr)
library(ggplot2)
library(scales)
library(foreach)
library(doParallel)
library(doFuture)
library(dplyr)
library(Rcpp)
library(RcppArmadillo)
# install.packages('RcppDist')
sourceCpp('rotavirus_normq.cpp')
source("choose_n_fits.R")
source("eval_logLik_pomp.R")

RUN_LEVEL = 1
NP_MIF       = switch(RUN_LEVEL, 5000, 50000)
NMIF         = switch(RUN_LEVEL, 4,    100)
ncores       = switch(RUN_LEVEL, 2,    36)
top_n_fits   = switch(RUN_LEVEL, 1,    12)
sim_rep      = switch(RUN_LEVEL, 1,    100) # Total number of simulations

#population number
pop <- 82372825
@

@@

<<OvOv-model, echo=False, massage=False, warning=False, results=hide>>=
# measurement model 
dmeas <- Csnippet("
                  if (ISNA(cases1)) {
                  lik = (give_log) ? 0 : 1;
                  } else {
                        lik =  dbinom(cases1, H1, q1, 1) +
                        dbinom(cases2, H2, q2, 1) +
                        dbinom(cases3, H3, q3, 1);

@@

# define parameters (without betas)
params_fixed <- c(gamma=1, delta1=1/(5*52),delta2=1/(55*52), alpha=1/(78.86912*52), 
                  mu=0, N=82372825, omega=1/(1*52))
# full-estimate-rinit
rinit <- Csnippet("
    double m = N/(S10+I10+R10+S20+I20+R20+S30+I30+R30);
    I1=nearbyint(m*I10);
    I2=nearbyint(m*I20);
    I3=nearbyint(m*I30);
    S1=nearbyint(m*S10);
    S2=nearbyint(m*S20);
    S3=nearbyint(m*S30);
    R1=nearbyint(m*R10);
    R2=nearbyint(m*R20);
    R3=nearbyint(m*R30);
    H1 = 0;
    H2 = 0;
    H3 = 0;
")

@@

# 
# sir_panel <- panelPomp::panelPomp(list(unit1=sir),
#                                   shared=NULL,
#                                   specific=wwr_mle |> 
#                                     as.matrix() |>
#                                     `colnames<-`("unit1")
# )
@

@@

<<maximization1, echo=False, massage=False, warning=False, results=hide>>=
set.seed(123)
sir_box <- rbind(
  beta1=c(10,15),
  beta2=c(0.2,0.4),
  beta3=c(0.3,0.5),
  phi=c(0.01,0.3),
  beta11=c(0.1,0.2),
  sigma_q=c(0.001,0.1),
  sigma_xi=c(65,70),
  S10 = c(0.01,0.05),
  I10 = c(0.0001, 0.0005),
  R10 = c(0.01, 0.05),
  S20 = c(0.01,0.7),
  I20 = c(0.000001,0.00002),
  R20 = c(0.001,0.005),
  S30 = c(0.01,0.5),
  I30 = c(0.000001, 0.0001),
  R30 = c(0.0001,0.005)
)

## Make it panelPomp
c(apply(sir_box,1,function(x)runif(1,min=x[1],max=x[2])),
  params_fixed) -> starting

bake(file = paste0("output/round_01/RUN_LEVEL_",RUN_LEVEL,"/ovov_mif_01.rds"),{ 
  registerDoFuture()
  plan(multicore, workers=ncores)
  mifs_global_01 <- foreach(i=1:ncores,.options.future=list(seed=652643293)) %dofuture% {
    pomp::mif2(
      sir,
      Np = NP_MIF,
      cooling.fraction.50 = 0.5,
      rw.sd = rw_sd(beta1=0.01,beta2=0.01,beta3=0.01,
                    beta11=0.01,phi=0.01,sigma_q=0.01,sigma_xi=0.01,
                    S10=ivp(0.24),I10=ivp(0.24),R10=ivp(0.24),
                    S20=ivp(0.24),I20=ivp(0.24),R20=ivp(0.24),
                    S30=ivp(0.24),I30=ivp(0.24),R30=ivp(0.24)),
      cooling.type = "geometric",
      Nmif = NMIF,
      params = starting,
      block = F
    ) 
  }
  mifs_global_01
})

bake(file = paste0("output/round_01/RUN_LEVEL_",RUN_LEVEL,"/ovov_el_01.rds"),{
  el_01 <- eval_logLik_pomp(mifs_global_01, ncores, NP_MIF, 4248930)
  el_01
})
@

@@

<<maximization2, echo=False, massage=False, warning=False, results=hide>>=
set.seed(1234)
sd_02           = 2/3

el_01 <- readRDS(paste0("output/round_01/RUN_LEVEL_",RUN_LEVEL,"/ovov_el_01.rds"))
starting_values_02 <- choose_n_fits(el_01, top_n_fits, ncores)

bake(file = paste0("output/round_02/RUN_LEVEL_",RUN_LEVEL,"/ovov_mif_02.rds"),{ 
  registerDoFuture()
  plan(multicore, workers=ncores)
  mifs_global_02 <- foreach(i=1:ncores,.options.future=list(seed=652643293)) %dofuture% {
    pomp::mif2(
      sir,
      Np = NP_MIF,
      cooling.fraction.50 = 0.5,
      rw.sd = rw_sd(beta1=0.01*sd_02,beta2=0.01*sd_02,beta3=0.01*sd_02,
                    beta11=0.01*sd_02,phi=0.01*sd_02,
                    sigma_q=0.01*sd_02,sigma_xi=0.01*sd_02,
                    S10=ivp(0.24)*sd_02,I10=ivp(0.24)*sd_02,R10=ivp(0.24)*sd_02,
                    S20=ivp(0.24)*sd_02,I20=ivp(0.24)*sd_02,R20=ivp(0.24)*sd_02,
                    S30=ivp(0.24)*sd_02,I30=ivp(0.24)*sd_02,R30=ivp(0.24)*sd_02),
      cooling.type = "geometric",
      Nmif = NMIF,
      params = starting,
      block = F
    ) 
  }
  mifs_global_02
})

bake(file = paste0("output/round_02/RUN_LEVEL_",RUN_LEVEL,"/ovov_el_02.rds"),{
  el_02 <- eval_logLik_pomp(mifs_global_02, ncores, NP_MIF, seed=4248930)
  el_02
})
@

@@

<<maximization3, echo=False, massage=False, warning=False, results=hide>>=
set.seed(12345)
sd_03           = (2/3)^2

el_02 <- readRDS(paste0("output/round_02/RUN_LEVEL_",RUN_LEVEL,"/ovov_el_02.rds"))
starting_values_03 <- choose_n_fits(el_02, top_n_fits, ncores)

bake(file = paste0("output/round_03/RUN_LEVEL_",RUN_LEVEL,"/ovov_mif_03.rds"),{ 
  registerDoFuture()
  plan(multicore, workers=ncores)
  mifs_global_03 <- foreach(i=1:ncores,.options.future=list(seed=652643293)) %dofuture% {
    pomp::mif2(
      sir,
      Np = NP_MIF,
      cooling.fraction.50 = 0.5,
      rw.sd = rw_sd(beta1=0.01*sd_03,beta2=0.01*sd_03,beta3=0.01*sd_03,
                    beta11=0.01*sd_03,phi=0.01*sd_03,
                    sigma_q=0.01*sd_03,sigma_xi=0.01*sd_03,
                    S10=ivp(0.24)*sd_03,I10=ivp(0.24)*sd_03,R10=ivp(0.24)*sd_03,
                    S20=ivp(0.24)*sd_03,I20=ivp(0.24)*sd_03,R20=ivp(0.24)*sd_03,
                    S30=ivp(0.24)*sd_03,I30=ivp(0.24)*sd_03,R30=ivp(0.24)*sd_03),
      cooling.type = "geometric",
      Nmif = NMIF,
      params = starting,
      block = F
    ) 
  }
  mifs_global_03
})

bake(file = paste0("output/round_03/RUN_LEVEL_",RUN_LEVEL,"/ovov_el_03.rds"),{
  el_03 <- eval_logLik_pomp(mifs_global_03, ncores, NP_MIF, seed=4248930)
  el_03
})
@

<<estimated-mle, echo=False, massage=False, warning=False>>=
mif_el <- readRDS(paste0("output/round_03/RUN_LEVEL_",RUN_LEVEL,"/ovov_el_03.rds")) 
# Can be changed if more rounds needed
est_mle_vec <- mif_el[which.max(mif_el$logLik),]
@

@@

<<pf-sim-100, echo=False, results=hide, massage=False, warning=False>>=
# Set WWR's MLE
wwr_mle_raw <- c(params_fixed, 
             "beta1" = 11.48,
             "beta2" = 0.25,
             "beta3" = 0.35,
             "phi" = 0.14,
             "beta11" = 0.16,
             "sigma_q" = 0.021,
             "sigma_xi" = 66.89)

i <- 1

bake(file=paste0("output/pfilter/ovov_sim_100_",RUN_LEVEL,".rds"),{
  ovov_sim_100 <- list()
  registerDoFuture()
  plan(multicore, workers=ncores)

  repeat {
    # Simulation
    sim1 <- simulate(t0 = 0,
                     times = c(1:416),
                     dmeasure = dmeas,
                     rmeasure = rmeas,
                     rprocess = discrete_time(rproc, delta.t = 1/4),
                     statenames = c("S1", "I1", "R1", "H1", 
                                    "S2", "I2", "R2", "H2",
                                    "S3", "I3", "R3", "H3",
                                    "q1", "q2", "q3"),
                     obsnames = c("cases1", "cases2", "cases3"),
                     paramnames = names(wwr_mle_raw),
                     accumvars = c("H1", "H2", "H3"),
                     rinit = rinit_wwr,
                     params = wwr_mle_raw)

    sim_data <- as.data.frame(sim1)
    if (!any(sim_data == 0)) {
      ovov_sim_100[[length(ovov_sim_100) + 1]] <- sim_data[, c("cases1", "cases2", "cases3")]
      i <- i + 1  # Increment only if the condition is satisfied
    } else {
      print(paste("Condition not met for iteration", i, "- retrying"))
    }

    if (length(ovov_sim_100) >= sim_rep) break  # Exit the loop when the desired number of successful iterations is reached
  }
  ovov_sim_100
})

# load("ovov_sim_100.Rdata")
@

<<pfilter_sim_100, echo=False, massage=False, warning=False, results=hide>>=
ovov_sim_100 <- readRDS(paste0("output/pfilter/ovov_sim_100_",RUN_LEVEL,".rds"))
bake(file=paste0("output/pfilter/pfilter_ovov_sim_100_",RUN_LEVEL,".rds"),{
  pfilter_ovov_sim_100 <- c()
  registerDoFuture()
  plan(multicore, workers=ncores)

  for(i in 1:length(ovov_sim_100)){
    dat <- data.frame(time=c(1:416), ovov_sim_100[[i]])
    dat |>
      pomp(t0=0,
           time="time",
           rprocess=discrete_time(rproc,delta.t=1/4), ## Obs time every week, but delta.t transition is actually everyday
           rinit=rinit_wwr,
           dmeasure=dmeas,
           rmeasure=rmeas,
           accumvars=c("H1","H2","H3"),
           statenames=c("S1","S2","S3",
                        "I1","I2","I3",
                        "R1","R2","R3",
                        "H1","H2","H3",
                        "q1","q2","q3"),
           paramnames = names(wwr_mle_raw)
      ) -> sir_wwr

    theta <- wwr_mle_raw

    foreach(i=1:ncores, .combine=c,
            .options.future=list(seed=998468235L)
    ) %dofuture% {
      pfilter(sir_wwr,Np=NP_MIF,save.states=T, params=theta)
    } -> pfs

    pfilter_ovov_sim_100[i] <-logmeanexp(logLik(pfs))  
  }
  pfilter_ovov_sim_100
})
@

<<palfilter_sim_100, echo=False, massage=False, warning=False, results=hide>>=
readRDS(paste0("output/pfilter/ovov_sim_100_",RUN_LEVEL,".rds"))
prop <- 420

init_dist <- c(3876549, 30351, 1315221, 57139612, 871, 302852, 19573727, 2550, 131092)
# initial_guess <- c(runif(1,5,20),runif(1,0.05,0.5),runif(1,0.2,1),runif(1,0.05,0.4),runif(1,0.05,0.4),runif(1,0.05,0.2),runif(1,10,100))
initial_guess <- c(11.48,0.25,0.35,0.14,0.16,0.021,66.89)

### Just a filter
Sys.time()

bake(file=paste0("output/pfilter/palfilter_ovov_100_",RUN_LEVEL,".Rdata"),{
  y <- c()
  registerDoFuture()
  plan(multicore, workers=ncores)
  for(i in 1:length(ovov_sim_100)){
    realdat <- ovov_sim_100[[i]] |> as.matrix()
    lik_list <- rotavirus_SMC_qropxi(init_dist,realdat , 9, 
                                     c(initial_guess[1],initial_guess[2],initial_guess[3],
                                       initial_guess[4],initial_guess[5]), 
                                     gamma_par = c(initial_guess[7],  1/initial_guess[7]), 
                                     norm_par = c(0.07,initial_guess[6]), 
                                     t(prop), NP_MIF,1)
    y[i] <- lik_list$log_lik
  }
  y
})

Sys.time()

# lik_list$ll_storage |> sum()

# x <- c()
# for(i in 1:10){
#   x[i] <- any(ovov_sim_100[[i]]==0)
# }
@

\begin{figure}[h]
\centering
\includegraphics[width=0.55\textwidth]{thesis-figs/PAL_vs_SMC_ovov_100.png}
\caption{\label{fig:PAL_vs_SMC_ovov_100}Log-likelihoods computed using two filtering methods for 100 randomly simulated datasets at the MLE of the OvOv model.
The particle filter and PAL used 50,000 particles.
The red line is the line $Lik(pfilter)=Lik(palfilter)$.
On average, the particle filter likelihood is $7.7$ log units higher than the PAL.
%%  -6224.245 - (-6231.97) = 7.725
}
\end{figure}

@@

\begin{table}[htbp]
  \centering
  \begin{threeparttable}
    \caption{Maximum likelihood estimation of parameters by PAL for three models \citep{wwr}, and by mif2 algorithm \citep{pomppackagepaper} for the OvOv model.
    \ed{LET'S DISCUSS WHETHER WE WANT TO SHOW JUST OV-OV HERE. IT WOULD BE SIMPLER.}
    }
    \label{tab:mlebywwr}
    \begin{tabular}{>{\raggedright\arraybackslash}p{2cm} *{3}{>{\centering\arraybackslash}p{1.5cm}} >{\centering\arraybackslash}p{2.5cm}}
      \toprule
      Parameter      & EqEq  & EqOv  & OvOv  & OvOv (\textbf{pomp})$^*$ \\
      \midrule
      \(\beta_1\)    & 12.15 & 12.74 & 11.48 & $\Sexpr{round(est_mle_vec["beta1"],2)}$\\
      \(\beta_2\)    & 0.22  & 0.21  & 0.25  & $\Sexpr{round(est_mle_vec["beta2"],2)}$\\
      \(\beta_3\)    & 0.34  & 0.31  & 0.35  & $\Sexpr{round(est_mle_vec["beta3"],2)}$\\
      \(\phi\)       & 0.017 & 0.14  & 0.14  & $\Sexpr{round(est_mle_vec["phi"],2)}$\\
      \(\rho\)       & 0.022 & 0.19  & 0.16  & $\Sexpr{round(est_mle_vec["beta11"],2)}$\\
      \(\sigma^2_q\) & n/a   & 0.042 & 0.021 & $\Sexpr{round(est_mle_vec["sigma_q"],2)}$\\
      \(\sigma_\xi\) & n/a   & n/a   & 66.89 & $\Sexpr{round(est_mle_vec["sigma_xi"],2)}$\\
      \addlinespace
      \multicolumn{1}{l}{\(S_{10}\)} & \multicolumn{3}{c}{$3876549^*$} & $\Sexpr{round(pop*est_mle_vec["S10"],0)}$\\
      \multicolumn{1}{l}{\(I_{10}\)} & \multicolumn{3}{c}{$30351$}     & $\Sexpr{round(pop*est_mle_vec["I10"],0)}$\\
      \multicolumn{1}{l}{\(R_{10}\)} & \multicolumn{3}{c}{$1315221$}   & $\Sexpr{round(pop*est_mle_vec["R10"],0)}$\\
      \multicolumn{1}{l}{\(S_{20}\)} & \multicolumn{3}{c}{$57139612$}  & $\Sexpr{round(pop*est_mle_vec["S20"],0)}$\\
      \multicolumn{1}{l}{\(I_{20}\)} & \multicolumn{3}{c}{$871$}       & $\Sexpr{round(pop*est_mle_vec["I20"],0)}$\\
      \multicolumn{1}{l}{\(R_{20}\)} & \multicolumn{3}{c}{$302852$}    & $\Sexpr{round(pop*est_mle_vec["R20"],0)}$\\
      \multicolumn{1}{l}{\(S_{30}\)} & \multicolumn{3}{c}{$19573727$}  & $\Sexpr{round(pop*est_mle_vec["S30"],0)}$\\
      \multicolumn{1}{l}{\(I_{30}\)} & \multicolumn{3}{c}{$2550$}      & $\Sexpr{round(pop*est_mle_vec["I30"],0)}$\\
      \multicolumn{1}{l}{\(R_{30}\)} & \multicolumn{3}{c}{$131092$}    & $\Sexpr{round(pop*est_mle_vec["R30"],0)}$\\
      \midrule
      \addlinespace
      \textbf{AIC}  & $98866.65$ & $15154.75$ & $13778.08$ & $\Sexpr{round((-2)*((length(wwr_mle) - length(params_fixed)) - est_mle_vec["logLik"]),2)}$\\
      \addlinespace
      \bottomrule
    \end{tabular}
    \begin{tablenotes}
      \small
      \item[*] The initial distribution parameter \(\lambda_0 = (S_{10}, I_{10}, R_{10}, S_{20}, I_{20}, R_{20}, S_{30}, I_{30}, R_{30})\) are assumed to be \textbf{fixed} in \cite{wwr}. The results of the estimates of the OvOv (\textbf{pomp}) model are obtained by using the Iterated Filtering algorithm with 3 rounds and 100 iterations, 50,000 particles, and 36 replicates in each round with the top 12 best fits in terms of likelihood chosen to be the starting value for the next round. 
    \end{tablenotes}
  \end{threeparttable}
\end{table}
